Generality in food webs scales with species richness but not with latitude

### 2015-06-19 manuscript/latitudepaper.tex
spellcheck

% \end{spacing}

@@

  %Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
  Food webs --networks of feeding links between species-- have been used for several decades to summarize the structure of 
  ecological communities~\citep{Petchey2008,Williams2000,Paine1966} and to understand how that structure relates to environmental variables 
  such as habitat type~\citep{Shurin2006,Briand1983}, primary productivity~\citep{Vermaat2009,Thompson2005a,Townsend1998}, and climate~\citep{Baiser2012,Petchey2010}. The latter
  variables in turn have strong gradients over latitude, with productivity and temperature both being higher in the tropics while 
  climate is more variable at high latitudes~\citep{Field1998}. These variables affect both the resources available and species' 
  metabolisms~\citep{Hechinger2011,White2011,OConnor2009,White2007}, and have been proposed as determinants of the strength of 
  interspecific interactions~\citep{Lang2012,Schleuning2012,Schemske2009}. 
  By modulating interactions between species, latitudinal gradients may also shape food-web structure.
  Indeed, these latitudinal environmental gradients have been put forward as potential drivers for the 
  latitudinal gradient in species richness, one of the most general and robust patterns in ecology~\citep{Schemske2009,Macpherson2002,Kaufman1995}.

@@

  If the latitude-niche breadth hypothesis is correct, there should also be direct relationships between latitude and the
  degree of specialization (i.e., the breadth of the Eltonian niche ~\citep{Elton1927}) of species within food webs.
  Attempts to unravel these effects, however,
  are complicated by known relationships between species richness and many other network properties~\citep{Riede2010}. 
  For example,
  narrower niches imply fewer links per species (i.e., greater specialization) in the tropics~\citep{Dyer2007,Marra1997}
  but see~\citep{Schleuning2012}). However,
  average numbers of links per species tend to increase in larger food webs~\citep{Dunne2006,Riede2010}. This means
  that a latitudinal effect on specialization may be obscured by a latitudinal gradient in species richness. If this is
  the case, it may still be possible to uncover effects of latitude on specialization by examining the shape of the scaling 
  relationship between specialization and species richness over changing latitude. Here, we use three measures of specialization:
  mean links per species, mean generality (number of prey), and mean vulnerability (number of predators). By testing whether
  latitude affects the scaling of each property with species richness, we test effects of latitude on specialization implied by
  the ``latitude-niche breadth hypothesis''.

@@

  Unlike species richness, only generality showed an overall increase with increasing
  proportions of basal resources in a web ($\beta_0$=-0.024, $p$=0.838;
  $\beta_0$=0.333, $p$=0.012; and $\beta_0$=-0.024, $p$=0.839 for link
  density, generality, and vulnerability respectively,
  Fig.~\ref{props_v_lat}). In each case, however, there were substantial
  effects of latitude on the scaling relationships. Thus, the scaling
  exponents for link density and vulnerability were negative for estaurine,
  lake, and stream communities and became more negative towards the poles
  ($\beta_{Latitude}$=-0.007, $p$=0.006 and $\beta_{Latitude}$=-0.007,
  $p$=0.006, respectively; Fig.~\ref{BIT}). For the scaling of link density and
  vulnerability in marine and terrestrial communities, the overall effect of
  latitude on scaling was positive ($\beta_{Latitude:Marine}$=0.009,
  $p$=0.047; $\beta_{Latitude:Terrestrial}$=0.013, $p$=0.002 and
  $\beta_{Latitude:Marine}$=0.009, $p$=0.047;
  $\beta_{Latitude:Terrestrial}$=0.013, $p$=0.002, respectively) such that
  there was little overall scaling relationship. In contrast, the scaling
  exponent of generality with the proportion of basal resources was positive
  near the equator and negative near the poles in estuarine, lake, and stream
  communities ($\beta_{Latitude}$=-0.009, $p$=0.002) while the scaling
  exponents for marine and terrestrial communities were not significantly
  different from zero ($\beta_{Latitude:Marine}$=0.011, $p$=0.028;
  $\beta_{Latitude:Terrestrial}$=0.015, $p$=0.001).

@@

  Unlike other ecosystem types, the scaling of generality in lake and stream
  food webs was stronger (i.e.,generality increased more steeply with
  increasing species richness) in higher-latitude food webs.In lake food webs,
  this trend was echoed in the scaling relationships between species richness
  and links per species and vulnerability. This means that species in tropical
  freshwater communities gain fewer additional feeding links per additional
  species in the web and that species in tropical lakes also gain fewer
  predators, and fewer links in general, per additional species than species
  in high-latitude lakes. These trends are consistent with the hypothesis that
  greater stability in the tropics leads to narrower
  niches~\citep{Vazquez2004} and a higher proportion of specialists. This may
  be partly due to high-latitude species tending to switch between different
  seasonally-available prey~\citep{Isaac2012,Wilhelm1999} while tropical
  freshwater ecosystems may have more stable composition.

  The scaling of link density, generality and vulnerability with proportions
  of species at different trophic levels, however, did not differ between lakes
  and most other food webs. Indeed, there was comparatively little variation in
  scaling with the trophic-level breakdown of a web across ecosystem types.
  Terrestrial food webs tended to exhibit less change in scaling over latitude
  than other ecosystem types, as did marine webs when considering scaling with
  the proportion of basal resources. More excitingly, the negative relationships
  between the proportion of top predators in a web and link density and
  vulnerability were weaker (i.e., the scaling exponent was more strongly
  negative) in low-latitude stream ecosystems. This may imply that high-latitude
  top predators tend to be more generalist than their low-latitude equivalents.
  More generally, the lack of correspondence between the scaling of food-web
  properties with species richness and with proportions of species at each
  trophic level suggests that the size and trophic breakdowns of a community can
  provide different information about the breadth of niches in that community.

@@

  Our results were inconsistent with the latitude-niche breadth hypothesis in estuarine, marine,
  and terrestrial communities but consistent with the hypothesis of greater specialization
  in the tropics in stream and lake food webs. This suggests that different mechanisms may structure
  food webs in different habitat types and that freshwater food webs in particular may be strongly
  affected by seasonal variation. In addition, different relationships between latitude and niche
  breadth in different habitat types goes some way towards explaining the equivocal support for
  the opposing hypotheses of narrower niches in the tropics~\citep{Vazquez2004} and broader 
  niche space in the tropics~\citep{Davies2007}; both have merit but appear
  to apply to different systems.

@@

\end{spacing}
\newpage

@@

\begin{figure}[h]
\centerline{\includegraphics*[height=.65\textheight]{Figures/by_TL/marginal/justS_marginal_latitude.eps}}
\caption{Changes to the scaling of link density with species richness across ecosystem
types and over latitude. We show the scaling exponent for species richness (black
line) with a 95\% confidence interval (in grey). Latitude is given in degrees from the equator
regardless of direction. The scaling relationships are based on N=163 empiricl food webs. 
Exponents for the scaling of generality and vulnerability with species richness were very 
similar to those for link density, except for the scaling of generality in streams where 
the size of the exponent increased towards the poles. See \emph{Fig. S2, Appendix 5} for 
all scaling relationships.}
\label{S}
\end{figure}

\begin{figure}[!H]
\centerline{\includegraphics*[height=.65\textheight]{Figures/by_TL/marginal/BIT_marginal_latitude.eps}}
\caption{Changes to the scaling of link density, generality, and vulnerability across ecosystem
types and over latitude. For each property we show the scaling exponent for the proportion of
basal resources (black line) with a 95\% confidence interval (in grey). Latitude is given in degrees
from the equator regardless of direction. The scaling relationships are based on N=163 empiricl food webs.
Exponents for the scaling relationships of generality and vulnerability with each proportion were very
similar to those of the scaling relationships with link density, except for the scaling of generality
with the proportion of top predators where there was no effect of latitude on the size of the exponent
in any ecosystem type. See \emph{Figs. S3-S5, Appendix 5} for all scaling relationships.}
\label{BIT}
\end{figure}

>>>>

\end{spacing}

@@

  %Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
  Food webs --networks of feeding links between species-- have been used for several decades to summarise the structure of 
  ecological communities~\citep{Petchey2008,Williams2000,Paine1966} and to understand how that structure relates to environmental variables 
  such as habitat type~\citep{Shurin2006,Briand1983}, primary productivity~\citep{Vermaat2009,Thompson2005a,Townsend1998}, and climate~\citep{Baiser2012,Petchey2010}. The latter
  variables in turn have strong gradients over latitude, with productivity and temperature both being higher in the tropics while 
  climate is more variable at high latitudes~\citep{Field1998}. These variables affect both the resources available and species' 
  metabolisms~\citep{Hechinger2011,White2011,OConnor2009,White2007}, and have been proposed as determinants of the strength of 
  interspecific interactions~\citep{Lang2012,Schleuning2012,Schemske2009}. 
  By modulating interactions between species, latitudinal gradients may also shape food-web structure.
  Indeed, these latitudinal environmental gradients have been put forward as potential drivers for the 
  latitudinal gradient in species richness, one of the most general and robust patterns in ecology~\citep{Schemske2009,Macpherson2002,Kaufman1995}.

@@

  If the latitude-niche breadth hypothesis is correct, there should also be direct relationships between latitude and the
  degree of specialisation (i.e., the breadth of the Eltonian niche ~\citep{Elton1927}) of species within food webs.
  Attempts to unravel these effects, however,
  are complicated by known relationships between species richness and many other network properties~\citep{Riede2010}. 
  For example,
  narrower niches imply fewer links per species (i.e., greater specialisation) in the tropics~\citep{Dyer2007,Marra1997}
  but see~\citep{Schleuning2012}). However,
  average numbers of links per species tend to increase in larger food webs~\citep{Dunne2006,Riede2010}. This means
  that a latitudinal effect on specialisation may be obscured by a latitudinal gradient in species richness. If this is
  the case, it may still be possible to uncover effects of latitude on specialisation by examining the shape of the scaling 
  relationship between specialisation and species richness over changing latitude. Here, we use three measures of specialisation:
  mean links per species, mean generality (number of prey), and mean vulnerability (number of predators). By testing whether
  latitude affects the scaling of each property with species richness, we test effects of latitude on specialisation implied by
  the ``latitude-niche breadth hypothesis''.

@@

  Unlike species richness, only generality showed an overall increase with increasing
  proportions of basal resources in a web ($\beta_0$=-0.024, $p$=0.838;
  $\beta_0$=0.333, $p$=0.012; and $\beta_0$=-0.024, $p$=0.839 for link
  density, generality, and vulnerability respectively,
  Fig.~\ref{props_v_lat}). In each case, however, there were substantial
  effects of latitude on the scaling relationships. Thus, the scaling
  exponents for link density and vulnerability were negative for estuarine,
  lake, and stream communities and became more negative towards the poles
  ($\beta_{Latitude}$=-0.007, $p$=0.006 and $\beta_{Latitude}$=-0.007,
  $p$=0.006, respectively; Fig.~\ref{BIT}). For the scaling of link density and
  vulnerability in marine and terrestrial communities, the overall effect of
  latitude on scaling was positive ($\beta_{Latitude:Marine}$=0.009,
  $p$=0.047; $\beta_{Latitude:Terrestrial}$=0.013, $p$=0.002 and
  $\beta_{Latitude:Marine}$=0.009, $p$=0.047;
  $\beta_{Latitude:Terrestrial}$=0.013, $p$=0.002, respectively) such that
  there was little overall scaling relationship. In contrast, the scaling
  exponent of generality with the proportion of basal resources was positive
  near the equator and negative near the poles in estuarine, lake, and stream
  communities ($\beta_{Latitude}$=-0.009, $p$=0.002) while the scaling
  exponents for marine and terrestrial communities were not significantly
  different from zero ($\beta_{Latitude:Marine}$=0.011, $p$=0.028;
  $\beta_{Latitude:Terrestrial}$=0.015, $p$=0.001).

@@

  Unlike other ecosystem types, the scaling of generality in lake and stream
  food webs was stronger (i.e.,generality increased more steeply with
  increasing species richness) in higher-latitude food webs. In lake food webs,
  this trend was echoed in the scaling relationships between species richness
  and links per species and vulnerability. This means that species in tropical
  freshwater communities gain fewer additional feeding links per additional
  species in the web and that species in tropical lakes also gain fewer
  predators, and fewer links in general, per additional species than species
  in high-latitude lakes. These trends are consistent with the hypothesis that
  greater stability in the tropics leads to narrower
  niches~\citep{Vazquez2004} and a higher proportion of specialists. This may
  be partly due to high-latitude species tending to switch between different
  seasonally-available prey~\citep{Isaac2012,Wilhelm1999} while tropical
  freshwater ecosystems may have more stable composition.

  The scaling of link density, generality and vulnerability with proportions
  of species at different trophic levels, however, did not differ between lakes
  and most other food webs. Indeed, there was comparatively little variation in
  scaling with the trophic-level breakdown of a web across ecosystem types.
  Terrestrial food webs tended to exhibit less change in scaling over latitude
  than other ecosystem types, as did marine webs when considering scaling with
  the proportion of basal resources. More excitingly, the negative relationships
  between the proportion of top predators in a web and link density and
  vulnerability were weaker (i.e., the scaling exponent was more strongly
  negative) in low-latitude stream ecosystems. This may imply that top predators
  in high-latitude streams tend to be more generalist than their low-latitude equivalents.
  More generally, the lack of correspondence between the scaling of food-web
  properties with species richness and with proportions of species at each
  trophic level suggests that the size and trophic breakdowns of a community can
  provide different information about the breadth of niches in that community.

@@

  Our results were inconsistent with the latitude-niche breadth hypothesis in estuarine, marine,
  and terrestrial communities but consistent with the hypothesis of greater specialisation
  in the tropics in stream and lake food webs. This suggests that different mechanisms may structure
  food webs in different habitat types and that freshwater food webs in particular may be strongly
  affected by seasonal variation. In addition, different relationships between latitude and niche
  breadth in different habitat types goes some way towards explaining the equivocal support for
  the opposing hypotheses of narrower niches in the tropics~\citep{Vazquez2004} and broader 
  niche space in the tropics~\citep{Davies2007}; both have merit but appear
  to apply to different systems.

@@

% \end{spacing}
\newpage

@@

\begin{figure}[h]
\centerline{\includegraphics*[height=.65\textheight]{Figures/by_TL/marginal/justS_marginal_latitude.eps}}
\caption{Changes to the scaling of link density with species richness across ecosystem
types and over latitude. We show the scaling exponent for species richness (black
line) with a 95\% confidence interval (in grey). Latitude is given in degrees from the equator
regardless of direction. The scaling relationships are based on N=163 empirical food webs. 
Exponents for the scaling of generality and vulnerability with species richness were very 
similar to those for link density, except for the scaling of generality in streams where 
the size of the exponent increased towards the poles. See \emph{Fig. S2, Appendix 5} for 
all scaling relationships.}
\label{S}
\end{figure}

\begin{figure}[!H]
\centerline{\includegraphics*[height=.65\textheight]{Figures/by_TL/marginal/BIT_marginal_latitude.eps}}
\caption{Changes to the scaling of link density, generality, and vulnerability across ecosystem
types and over latitude. For each property we show the scaling exponent for the proportion of
basal resources (black line) with a 95\% confidence interval (in grey). Latitude is given in degrees
from the equator regardless of direction. The scaling relationships are based on N=163 empirical food webs.
Exponents for the scaling relationships of generality and vulnerability with each proportion were very
similar to those of the scaling relationships with link density, except for the scaling of generality
with the proportion of top predators where there was no effect of latitude on the size of the exponent
in any ecosystem type. See \emph{Figs. S3-S5, Appendix 5} for all scaling relationships.}
\label{BIT}
\end{figure}
